---
title: "Assignment-3: Complex Queries using RStudio/SQLite"
author: "Dhanashree"
date: "15/03/2020"
output: html_document
---

```{r setup, include=FALSE}
library("RSQLite")
library("knitr")
library("DBI")
dbLGSales=dbConnect(RSQLite::SQLite(),dbname="~/Desktop/ITC6000/RSQLite/databases/LG_Sales.db")
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(connection = dbLGSales)
```

```{r}
dbListTables(dbLGSales)
```
# Question-1: List all customers using a negative subquery who do not reside in the New England
region {if needed, do a google search to find out what states are considered part of New England region}. Show
customer’s full name, city, state, and zip in the sorted order. 
```{sql connection=dbLGSales, max.print=-1}
SELECT * 
FROM   lgcustomer 
WHERE  cust_code IN (SELECT cust_code 
                         FROM   lgcustomer 
                         WHERE  cust_state NOT IN ( 'ME', 'NH', 'MA', 'RI', 'CT'  )) 
ORDER  BY cust_code 
```

#Question-2: Generate inventory status report of all products that have current inventory level (i.e.,Prod_QOH) lower than 120% of Reorder points (i.e., Prod_Min). Include a column for Procurement Instruction indicating “order immediately” if current inventory is not above the reorder point; and as “order soon” if current inventory is above reorder point but lesser than 120% of reorder pint. Final query output should show Product SKU, Brand Name, Brand Type, Product Type, Product Base, Product Category, Product Price, Quantity On Hand, Reorder level, Procurement Instruction, Vendor Name
```{sql connection=dbLGSales, max.print=-1}
SELECT LP.prod_sku   'Product SKU', 
       brand_name    'Brand Name', 
       brand_type    'Brand Type', 
       prod_type     'Product Type', 
       prod_base     'Product Base', 
       prod_category 'Product Category', 
       prod_price    'Product Price', 
       prod_qoh      'Quantity On Hand', 
       prod_min      'Reorder level', 
       CASE 
         WHEN prod_qoh <= prod_min THEN 'Order Immediately' 
         WHEN ( ( prod_qoh > prod_min ) 
                AND ( prod_qoh < ( 1.2 * prod_min ) ) ) THEN 'Order Soon' 
         ELSE 'Sufficient Stock in Inventory' 
       END           'Procurement Instruction', 
       vend_name     'Vendor Name' 
FROM   lgproduct LP 
       INNER JOIN lgbrand LB 
               ON LP.brand_id = LB.brand_id 
       INNER JOIN lgsupplies LS 
               ON LS.prod_sku = LP.prod_sku 
       INNER JOIN lgvendor LV 
               ON LV.vend_id = LS.vend_id 
WHERE  prod_qoh < ( 1.2 * prod_min ) 
```

#Question-3: Develop a query to generate necessary information for monthly sales (in $$) report of paint products for each brand name of different brand types. The report should show $0 if no product was sold in a month. Show the results in the sorted order of brand types and brand names.
```{sql connection=dbLGSales, max.print=-1}
WITH months 
     AS (SELECT "01", 
                "2015" 
         UNION 
         SELECT "02", 
                "2015" 
         UNION 
         SELECT "03", 
                "2015" 
         UNION 
         SELECT "04", 
                "2015" 
         UNION 
         SELECT "05", 
                "2015" 
         UNION 
         SELECT "06", 
                "2015" 
         UNION 
         SELECT "07", 
                "2015" 
         UNION 
         SELECT "08", 
                "2015" 
         UNION 
         SELECT "09", 
                "2015" 
         UNION 
         SELECT "10", 
                '2015' 
         UNION 
         SELECT "11", 
                "2015" 
         UNION 
         SELECT "12", 
                "2015" 
         UNION 
         SELECT "01", 
                "2016"), 
     brands 
     AS (SELECT brand_name, 
                brand_type 
         FROM   lgbrand) 
         
SELECT view1.*, 
       printf("$ %.2f", Ifnull(monthlysales, 0)) MonthlySales
FROM   (SELECT B.brand_name, 
               B.brand_type, 
               M."01"   AS Month, 
               M."2015" AS Year 
        FROM   brands B 
               CROSS JOIN months M) view1 
       LEFT JOIN  (SELECT LB.brand_name, 
                                 LB.brand_type, 
                                 Strftime("%Y", LI.inv_date) AS Year, 
                                 Strftime("%m", LI.inv_date) AS Month, 
                              Sum(Ifnull((line_qty*line_Price), 0)) 'MonthlySales'    
                          FROM   lgbrand LB 
                                 INNER JOIN lgproduct LP 
                                        ON LB.brand_id = LP.brand_id 
                                 INNER JOIN lgline LL 
                                        ON LL.prod_sku = LP.prod_sku 
                                 INNER JOIN lginvoice LI 
                                        ON LI.inv_num = LL.inv_num
                                        GROUP  BY brand_name, 
                            brand_type, 
                            year, 
                            month) 
                                         view2 
              ON view1.brand_name = view2.brand_name 
                 AND view1.brand_type = view2.brand_type 
                 AND view1.year = view2.year 
                 AND view1.month = view2.month 
ORDER  BY 1, 
          4, 
          3 
```

# Question-4: Generate a list of lowest and highest selling paint products for each category of paint products in each state during each quarter of 2015. Display all necessary columns.
```{sql connection=dbLGSales, max.print=-1}
WITH s AS 
( 
         SELECT   state, 
                  year, 
                  CASE 
                           WHEN months < "04" THEN "Q1" 
                           WHEN months < "07" THEN "Q2" 
                           WHEN months < "10" THEN "Q3" 
                           ELSE "Q4" 
                  END AS quarters, 
                  prod_category, 
                  prod_sku, 
                  productsales , 
                  row_number() OVER ( partition BY state, year, 
                  CASE 
                           WHEN months < "04" THEN "Q1" 
                           WHEN months < "07" THEN "Q2" 
                           WHEN months < "10" THEN "Q3" 
                           ELSE "Q4" 
                  END , prod_category ORDER BY productsales ASC) AS rnk1, 
                  row_number() OVER ( partition BY state, year, 
                  CASE 
                           WHEN months < "04" THEN "Q1" 
                           WHEN months < "07" THEN "Q2" 
                           WHEN months < "10" THEN "Q3" 
                           ELSE "Q4" 
                  END , prod_category ORDER BY productsales ASC) AS rnk2
FROM   ( 
                  SELECT     lc.cust_state state, 
                             lp.prod_category, 
                             lp.prod_sku, 
                             Strftime("%Y", li.inv_date) year, 
                             Strftime("%m", li.inv_date) months, 
                             Sum(line_qty*line_price)    productsales 
                  FROM       lgproduct LP 
                  INNER JOIN lgline LL 
                  ON         ll.prod_sku=lp.prod_sku 
                  INNER JOIN lginvoice LI 
                  ON         li.inv_num=ll.inv_num 
                  INNER JOIN lgcustomer LC 
                  ON         lc.cust_code=li.cust_code 
                  WHERE      year="2015" 
                  GROUP BY   state, 
                             months, 
                             lp.prod_category, 
                             lp.prod_sku 
                  ORDER BY   1, 
                             5, 
                             2, 
                             3))

SELECT view2.state 'State', 
       view2.year 'Year', 
       view2.quarters 'Quarter', 
       view2.prod_category 'Product Category', 
       view1.prod_sku      'Product with Minimum Sales', 
       view2.prod_sku      'Product with Maximum Sales' 
FROM   (SELECT state, 
               year, 
               quarters, 
               prod_category, 
               prod_sku, 
               Min(rnk1) AS MinSalesProduct 
        FROM   s 
        GROUP  BY state, 
                  year, 
                  quarters, 
                  prod_category)view1 
       INNER JOIN (SELECT state, 
                          year, 
                          quarters, 
                          prod_category, 
                          prod_sku, 
                          Max(rnk2) AS MaxSalesProduct 
                   FROM   s 
                   GROUP  BY state, 
                             year, 
                             quarters, 
                             prod_category)view2 
               ON view1.state = view2.state 
                  AND view1.year = view2.year 
                  AND view1.quarters = view2.quarters 
                  AND view1.prod_category = view2.prod_category 
ORDER  BY 1, 
          2, 
          3, 
          4  
``` 
  
# Question-5: Develop a query to generate annual sales variance report of sales for each paint product (Prod_SKU) within each brand type, product type, product base, product category. The sales variance for a product is calculated as the deviations from (a) national average annual sales, and (b) state average annual sales for all paint products within their group (identified thru the combination of brand type, product type, product base, and product category). The result should show brand type, product type, product base, product category, product SKU, aggregation level, annual average sales, sales variance. The column “aggregation level” will be either ‘national’ or state codes.
```{sql connection=dbLGSales, max.print=-1}
WITH nationalsales 
     AS (SELECT Strftime("%Y", LI.inv_date)                year, 
                LB.brand_type, 
                LP.prod_type, 
                LP.prod_base, 
                LP.prod_category, 
                Sum(line_qty * line_price) / Sum(line_qty) NationalAvgSales
         FROM   lgproduct LP 
                INNER JOIN lgbrand LB 
                        ON LP.brand_id = LB.brand_id 
                INNER JOIN lgline LL 
                        ON LP.prod_sku = LL.prod_sku 
                INNER JOIN lginvoice LI 
                        ON LL.inv_num = LI.inv_num 
         GROUP  BY year, 
                   LB.brand_type, 
                   LP.prod_type, 
                   LP.prod_base, 
                   LP.prod_category), 
     statessales 
     AS (SELECT Strftime("%Y", LI.inv_date)                year, 
                LC.cust_state                              STATE, 
                LB.brand_type, 
                LP.prod_type, 
                LP.prod_base, 
                LP.prod_category, 
                Sum(line_qty * line_price) / Sum(line_qty) StateAvgSales 
         FROM   lgproduct LP 
                INNER JOIN lgbrand LB 
                        ON LP.brand_id = LB.brand_id 
                INNER JOIN lgline LL 
                        ON LP.prod_sku = LL.prod_sku 
                INNER JOIN lginvoice LI 
                        ON LL.inv_num = LI.inv_num 
                INNER JOIN lgcustomer LC 
                        ON LC.cust_code = LI.cust_code 
         GROUP  BY year, 
                   state, 
                   LB.brand_type, 
                   LP.prod_type, 
                   LP.prod_base, 
                   LP.prod_category) 

SELECT year, 
       brand_type, 
       prod_type, 
       prod_base, 
       prod_category, 
       prod_sku, 
       'National' AggregationLevel, 
       AnnualAverageSales, 
       Printf("%f", Sum(variance)) 'SalesVariance' 
FROM   (SELECT Strftime("%Y", li.inv_date)                           year, 
               lb.brand_type, 
               lp.prod_type, 
               lp.prod_base, 
               lp.prod_category, 
               lp.prod_sku, 
               ll.line_qty                                           Qty, 
               ll.line_price                                         Price, 
               NS.nationalavgsales 'AnnualAverageSales', 
               ( ll.line_price - NS.nationalavgsales ) * ll.line_qty 'Variance'
        FROM   lgproduct LP 
               INNER JOIN lgbrand LB 
                       ON lp.brand_id = lb.brand_id 
               INNER JOIN lgline LL 
                       ON lp.prod_sku = ll.prod_sku 
               INNER JOIN lginvoice LI 
                       ON ll.inv_num = li.inv_num 
               INNER JOIN nationalsales NS 
                       ON NS.year = Strftime("%Y", li.inv_date) 
                          AND NS.brand_type = lb.brand_type 
                          AND NS.prod_type = Lp.prod_type 
                          AND NS.prod_base = lp.prod_base 
                          AND NS.prod_category = LP.prod_category) 
GROUP  BY year, 
          brand_type, 
          prod_type, 
          prod_base, 
          prod_category, 
          prod_sku 
UNION ALL 
SELECT year, 
       brand_type, 
       prod_type, 
       prod_base, 
       prod_category, 
       prod_sku, 
       cust_state 'AggregationLevel',
       AnnualAverageSales, 
       Printf("%f", Sum(variance)) 'SalesVariance' 
FROM   (SELECT LC.cust_state, 
               Strftime("%Y", li.inv_date)                        year, 
               lb.brand_type, 
               lp.prod_type, 
               lp.prod_base, 
               lp.prod_category, 
               lp.prod_sku, 
               ll.line_qty                                        Qty, 
               ll.line_price                                      Price, 
               SS.stateavgsales 'AnnualAverageSales', 
               ( ll.line_price - SS.stateavgsales ) * ll.line_qty 'Variance'
        FROM   lgproduct LP 
               INNER JOIN lgbrand LB 
                       ON lp.brand_id = lb.brand_id 
               INNER JOIN lgline LL 
                       ON lp.prod_sku = ll.prod_sku 
               INNER JOIN lginvoice LI 
                       ON ll.inv_num = li.inv_num 
               INNER JOIN lgcustomer LC 
                       ON LC.cust_code = LI.cust_code 
               INNER JOIN statessales SS 
                       ON SS.year = Strftime("%Y", li.inv_date) 
                          AND SS.state= LC.Cust_State
                          AND SS.brand_type = lb.brand_type 
                          AND SS.prod_type = Lp.prod_type 
                          AND SS.prod_base = lp.prod_base 
                          AND SS.prod_category = LP.prod_category) 
GROUP  BY cust_state, 
          year, 
          brand_type, 
          prod_type, 
          prod_base, 
          prod_category, 
          prod_sku
ORDER BY 1,7,2,3,4,5          
```

